Fast Sorting on Modern Computers

### 2005-08-14 docs/tech_report/conclusions.tex
renamed cases of 2[ \-]level adaptive to two-level

We found that algorithm S has a lower instruction count than algorithm N, but
that this does not necessarily translate to better performance. Our results show
also that the level 1 cache needs to be taken into account, that the extra
instruction in multi-mergesort make it slower than double-aligned tiled
mergesort, and that multi-mergesort has regular branch predictive access
patterns during the k-way merge, which can be taken advantage of by a 2-level
adaptive predictor.

@@

Using a k-way merge, multi-mergesort had a great number of extra branch
predictions. While many of these were mispredicted using the bimodal predictor,
the 2 level adaptive predictor was able to predict a higher
proportion correctly.

>>>>

We found that algorithm S has a lower instruction count than algorithm N, but
that this does not necessarily translate to better performance. Our results show
also that the level 1 cache needs to be taken into account, that the extra
instruction in multi-mergesort make it slower than double-aligned tiled
mergesort, and that multi-mergesort has regular branch predictive access
patterns during the k-way merge, which can be taken advantage of by a two-level
adaptive predictor.

@@

Using a k-way merge, multi-mergesort had a great number of extra branch
predictions. While many of these were mispredicted using the bimodal predictor,
the two-level adaptive predictor was able to predict a higher
proportion correctly.

### 2005-08-14 docs/tech_report/method.tex
fixed usage of '32 byte' without a hyphen
renamed cases of 2[ \-]level adaptive to two-level

The SimpleScalar settings were chosen to be similar to LaMarca's simulations.
LaMarca used the \n{Atom} simulator, which was used to test software for the DEC
AlphaStation. It simulated a 2MB direct-mapped cache with a 32 byte cache line.
Other settings were also chosen to correspond to the DEC Alpha: 8KB level 1 data
cache, 8KB level 1 instruction cache and shared instruction and data level 2
cache.

@@

We chose a similar approach for our branch predictor caches. We aimed to
simulate bimodal predictors as well as 2 level adaptive, in order to be able to
compare them together. We simulated bimodal predictors with a 512, 1024 and
2048 byte table, and 2 level adaptive predictors with 512, 1024, 2048 and 4096
entries.

>>>>

The SimpleScalar settings were chosen to be similar to LaMarca's simulations.
LaMarca used the \n{Atom} simulator, which was used to test software for the DEC
AlphaStation. It simulated a 2MB direct-mapped cache with a 32-byte cache line.
Other settings were also chosen to correspond to the DEC Alpha: 8KB level 1 data
cache, 8KB level 1 instruction cache and shared instruction and data level 2
cache.

@@

We chose a similar approach for our branch predictor caches. We aimed to
simulate bimodal predictors as well as two-level adaptive, in order to be able to
compare them together. We simulated bimodal predictors with a 512, 1024 and
2048 byte table, and two-level adaptive predictors with 512, 1024, 2048 and 4096
entries.

### 2005-08-14 docs/tech_report/quick.tex
renamed cases of 2[ \-]level adaptive to two-level

The branch prediction results are shown in Figure \ref{quick branch misses}. As
expected, there was a a dramatic increase in the binary search multi-quicksort
branch misses, but only when using a bimodal predictor. When using a 2-level
adaptive predictor, the number of misses is, in general, slightly higher, but it
does not spike due to the binary search. While it performs worse than the
bimodal predictor, it appears that 2-level adaptive predictors can adapt to the
erratic branches of a binary search.

>>>>

The branch prediction results are shown in Figure \ref{quick branch misses}. As
expected, there was a a dramatic increase in the binary search multi-quicksort
branch misses, but only when using a bimodal predictor. When using a two-level
adaptive predictor, the number of misses is, in general, slightly higher, but it
does not spike due to the binary search. While it performs worse than the
bimodal predictor, it appears that two-level adaptive predictors can adapt to
the erratic branches of a binary search.
